Latent Variable Clustering for Classification of Events in the AT-TPC

### 2020-04-03 doc/Manuscript/article.tex
Updates from Overleaf
Experimental details

Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not practical nowadays because of the large quantities of data. Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently. In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods. The classification and clustering of events based on various ML algorithm is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.

\section{Experimental Details} 
(46Ar)
\section{Methods}\label{sec:methods}
\subsection{Why machine learning}

@@

\begin{itemize}
    \item Reiterate the aim of unsupervised  clustering of events.
    \item Challenges from a machine learning perspective. 
    \begin{itemize}
        \item Supervised vs. Unsupervised learning
        \item Traditional unupervised - distances in high dimensional spaces etc. 
    \end{itemize}
    \item Digression here to explain some ML concepts? Link back to Conv part of Michelles paper? 
    \item Opportunities presented by: 
    \begin{itemize}
        \item Transfer learning (train on VGG use on AT-TPC)
        \item Compression is understanding (leverage autoencoders) 
    \end{itemize}
    \item Building on Kuchera et. al we investigate the VGG16 networks application. 
    \item Novel contribution by applying clustering autoencoder networks. 
\end{itemize}

>>>>

Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not practical nowadays because of the large quantities of data. Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently. In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods. The classification and clustering of events based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.

\section{Experimental Details} 
The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (also called excitation function), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar. Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018}. 
The experiment was performed at the National Superconducting Cyclotron Facility (NSCL) where a $^{46}$Ar beam was produced via fragmentation of a $^{48}$Ca beam on a $^9$Be target at about 140 MeV/u. The $^{46}$Ar isotopes were then filtered, thermalized, and finally re-accelerated to 4.6 MeV/u by a linear accelerator. This scheme was used to produce a low emittance beam necessary to guarantee a good energy resolution in the excitation function. Because of the energy loss of the $^{46}$Ar beam particles as they traverse the target gas volume, the position of the reaction vertex along the beam axis is directly related to the energy at which the reaction occurs. This allows the AT-TPC to measure the excitation function over a wide range of energies from a single beam energy.
\section{Methods}\label{sec:methods}
\subsection{Why machine learning}

@@

\begin{itemize}
    \item Reiterate the aim of unsupervised  clustering of events.
    \item Challenges from a machine learning perspective. 
    \begin{itemize}
        \item Supervised vs. Unsupervised learning
        \item Define Classifcation and Clustering 
        \item Define latent space (refer)
        \item Traditional unupervised - distances in high dimensional spaces etc. 
    \end{itemize}
    \item Digression here to explain some ML concepts? Link back to Conv part of Michelles paper? 
    \item Opportunities presented by: 
    \begin{itemize}
        \item Transfer learning (train on VGG use on AT-TPC)
        \item Compression is understanding (leverage autoencoders) 
    \end{itemize}
    \item Building on Kuchera et. al we investigate the VGG16 networks application. 
    \item Novel contribution by applying clustering autoencoder networks. 
\end{itemize}
